---
title: "MEE_reproduce_2"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEE_reproduce_2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

```{r setup}
library(sizeSpectra)
```

# Reproduce Figures 3, ....TODO.... of MEE paper

This vignette reproduces figures and tables in TODO: ref paper. See the
paper for details, the aim of the vignette is to enable users to reproduce the
results so that they can see what functions are appropriate for their own data,
and perform further analyses as required.

COPIES FIRST ONE FOR NOW (as template)

```{r}
library(sizeSpectra)
```

## Figure 3 - Histograms of estimated exponent *b* for simulated data

Fit histograms of 10,000 simulated data sets each with .....

The fitting takes a while, so results are saved in `eight.results.default`. To
run from scratch use the code in `data-raw/simulate-data.R`.
```{r, fig.height=7}

list2env(eight.results.default, envir=.GlobalEnv)    # Extract previously saved
                                        # results. To obtain from scratch see data-raw/simulate-data.R

# Plotting
# brange = range( c( Llin.rep, LT.rep, LTplus1.rep, LBmiz.rep,
#    LBbiom.rep, LBNbiom.rep, MLE.rep))
# Comes out as -3.38, 0.233 for seed=42, num.reps = 10000
xrange = c(-3.5, 0.5)         # range of x-axis for histograms - actually to define bins
xbigticks = seq(-3.5, 0.5, 0.5) # NOW THE DEFAULT

xsmallticks = seq(xrange[1], xrange[2], by=0.1)  # ALSO now default
breakshist = seq(xrange[1], xrange[2], by=3/61)
         # bin ends at 0, 2 is centered. width of 4/81 from solving
         #  the mean of a bin equals 2, ((N+1)w + Nw)/2 = 2, and setting number
         #  of the bin that includes 2, N, =40 (which is how many you get with
         #  0.05, which looked good).

# figheight = 7 # 5.6
# figwidth = 5.7    # 5.7 inches for JAE
#postscript("fitting3rep.eps", height = figheight, width = figwidth,
#           horizontal=FALSE,  paper="special")
par(omi = c(0.12, 0.05, 0.12, 0.0))      # outer margins in inches
par(mfrow=c(4,2)) #7,1))

oldmai = par("mai")
par(mai=c(0.5, 0.5, 0.1, 0.3)) # Affects all figures if don't change agaiin
par(xaxs="i", yaxs="i")    # Have to define here for hist
par(mgp=c(2.0, 0.5, 0))    # puts axes labels closer I think
par(cex = 0.8)             # With no option all text comes out a bit small

vertCol = "red"            # Colour for vertical lines
vertThick = 1              # Thickness for vertical lines

xLim = c(-3.2, -1.5)
xLimLlin = xLim + 2
ylimA = c(0, 5500)
ylimLlin = c(0, 10000)
cexAxis = 0.9      # font size for axes labels to make the y ones fit okay

# Llin.rep has different breakhist so that 0 is a breakpoint. Same width as
#  others, just shifted.
hist(Llin.rep.df$slope, xlim=xLimLlin, breaks=breakshist - breakshist[31],
  xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimLlin)  #  ylim=c(0,1100),
axis(1, at=xbigticks, labels = xbigticks, mgp=c(1.7,0.7,0), cex.axis=cexAxis)  # big ticks
axis(1, at=xsmallticks, labels=rep("",length(xsmallticks)), tcl=-0.2)
axis(2, at=c(0, 5000, 10000),
         labels = c(0, 5000, 10000),
         mgp=c(1.7,0.7,0), cex.axis=cexAxis)  # big ticks labelled
axis(2, at=seq(0, 10000, 1000),
         labels = rep("", 11),
         mgp=c(1.7,0.7,0))  # big ticks unlabelled
abline(v=b.known, col=vertCol, lwd=vertThick)
inset = c(-0.08, -0.08)     # inset distance of legend
legend("topleft", "(a) Llin", bty="n", inset=inset)

# TODO think these three lines can be deleted
figlabpos = 0.93    # proportion of x and y axis lengths to put (a) in etc.
xlabpos = 0.75       # just play with a number, as now using pos=4 in text
text( xlabpos, figlabpos * 1100, "(a) Llin", pos=4)

hist(LT.rep.df$slope, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(b) LT", bty="n", inset=inset)

hist(LTplus1.rep.df$slope, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(c) LTplus1", bty="n", inset=inset)

hist(LBmiz.rep.df$slope - 1, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(d) LBmiz", bty="n", inset=inset)

hist(LBbiom.rep.df$slope - 2, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(e) LBbiom", bty="n", inset=inset)

hist(LBNbiom.rep.df$slope - 1, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(f) LBNbiom", bty="n", inset=inset)


hist(LCD.rep.df$slope - 1, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(g) LCD", bty="n", inset=inset)

hist(MLE.rep.df$b, xlim=xLim, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = ylimA)  #  ylim=c(0,1100),
histAxes()
legend("topleft", "(h) MLE", bty="n", inset=inset)

# text(-2.5, 0, expression(paste("Estimate of ", italic(b))))
mtext(expression(paste("Estimate of ", italic(b))), side=1, outer=TRUE, line=-1)
# mtext("hello", side=1, outer=TRUE, line=-1)
```

Figure A.3 for just the MLE and MLEfix methods:

```{r}

par(omi = c(0.12, 0.05, 0.12, 0.0))      # outer margins in inches
par(mfrow=c(2,1))

par(mai=c(0.5, 0.5, 0.1, 0.3))
par(xaxs="i", yaxs="i")    # Have to define here for hist
par(mgp=c(2.0, 0.5, 0))    # puts axes labels closer I think
par(cex = 0.8)             # With no option all text comes out a bit small

xLimFix = c(-2.5, -1.5)
yLimFix = c(0, 6000)
hist(MLE.rep.df$b, xlim=xLimFix, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = yLimFix)  #  ylim=c(0,1100),
histAxes()
axis(2, at=6000, labels = 6000, mgp=c(1.7,0.7,0), cex.axis=cexAxis)
                                        # big tick label
legend("topleft", "(a) MLE", bty="n", inset=0.5*inset)

hist(MLEfix.rep.df$b, xlim=xLimFix, breaks=breakshist, xlab="", ylab="Frequency",
  main="", axes=FALSE, ylim = yLimFix)  #  ylim=c(0,1100),
histAxes()
axis(2, at=6000, labels = 6000, mgp=c(1.7,0.7,0), cex.axis=cexAxis)
                                        # big tick label
legend("topleft", "(b) MLEfix", bty="n", inset=0.5*inset)

mtext(expression(paste("Estimate of ", italic(b))), side=1, outer=TRUE, line=-1)
```

Table 2, summarising the results for each method:

<!-- For Latex table for each row, see commented out code below. -->
<!-- I had that code writen already, so adapted it to make an -->
<!-- Rmarkdown one for the vignette (or could have made a data frame and used -->
<!-- kable.
<!-- Doing table in terms of b, not slope, so have to subtract 1 or 2 from some results.
-->

|Method | Slope represents | 5\% quantile |Median  |Mean  |95\% quantile |Percentage below true|
|:----- | :--------------- | -----------: |-----:  |---:  |-----------:  |--------------------:|
|Llin   | -                | `r qqtab(Llin.rep.df$slope, dig=2, quants=c(0.05, 0.95))` |
|LT     | *b*              | `r qqtab(LT.rep.df$slope, quants=c(0.05, 0.95))` |
|LTplus1| *b*              | `r qqtab(LTplus1.rep.df$slope, quants=c(0.05, 0.95))` |
|LBmiz  | *b+1*            | `r qqtab(LBmiz.rep.df$slope - 1, quants=c(0.05, 0.95))` |
|LBbiom | *b+2*            | `r qqtab(LBbiom.rep.df$slope - 2, quants=c(0.05, 0.95))` |
|LBNbiom| *b+1*            | `r qqtab(LBNbiom.rep.df$slope - 1, quants=c(0.05, 0.95))` |
|LCD    | *b+1*            | `r qqtab(LCD.rep.df$slope - 1, quants=c(0.05, 0.95))` |
|MLE    | *b*              | `r qqtab(MLE.rep.df$b, quants=c(0.05, 0.95))` |

For MLEbin method see next vignette.TODO. And create data set with 10,000 reps.


<!--

# Prints Latex code for table that summarises the results
# Going to do in terms of b, not slope, so subtract 1 or 2 from some results:

# 5% and 95% values:
print("Method & Slope represents & 5% quantile & Median & Mean & 95% quantile & Percentage below true \\")
print(paste("Llin & - & ", qqtab(Llin.rep, dig=2, quants=c(0.05, 0.95)), "\\"))
print(paste("LT & $b$ & ", qqtab(LT.rep, quants=c(0.05, 0.95)), "\\"))
print(paste("LTplus1 & $b$ & ", qqtab(LTplus1.rep, quants=c(0.05, 0.95)), "\\"))
print(paste("LBmiz & $b+1$ & ", qqtab(LBmiz.rep - 1, quants=c(0.05, 0.95)), "\\"))
print(paste("LBbiom & $b+2$ & ", qqtab(LBbiom.rep - 2, quants=c(0.05, 0.95)), "\\"))
print(paste("LBNbiom & $b+1$ & ", qqtab(LBNbiom.rep - 1, quants=c(0.05, 0.95)), "\\"))
print(paste("LCD & $b+1$ & ", qqtab(LCD.rep - 1, quants=c(0.05, 0.95)), "\\"))
print(paste("MLE & $b$ & ", qqtab(MLE.rep, quants=c(0.05, 0.95)), "\\hline"))
print(paste("% MLEfix & $b$ & ", qqtab(MLEfix.rep, quants=c(0.05, 0.95)), "\\"))

# For slopes:
#print(paste("LBmiz &  ", qqtab(LBmiz.rep, true=b.known+1, quants=c(0.05, 0.95)), "\\"))
#print(paste("LBbiom &  ", qqtab(LBbiom.rep, true=b.known+2, quants=c(0.05, 0.95)), "\\"))
#print(paste("LBNbiom & ", qqtab(LBNbiom.rep, true=b.known+1, quants=c(0.05, 0.95)), "\\"))
#print(paste("LCD & ", qqtab(LCD.rep, true=b.known+1, quants=c(0.05, 0.95)), "\\"))

save.image(file = "fitting3rep.RData")


-->

# From fitting3conf.r

```{r, fig.height=7}
par(omi = c(0.14, 0, 0.1, 0.15))      # outer margins in inches
par(mfrow=c(4,2)) #7,1))

oldmai = par("mai")    #   0.6732 0.5412 0.5412 0.2772  inches I think,
                       #    may be indpt of fig size
par(mai=c(0.3, 0.5, 0.08, 0))  # Affects all four figures if don't change agaiin
par(xaxs="i", yaxs="i")    # Have to define here for hist
par(mgp=c(2.0, 0.5, 0))    # puts axes labels closer I think
par(cex = 0.8)             # With no option all text comes out a bit small
vertThick = 1              # Thickness for vertical lines

# Each of these plots a panel for one method. Define xLim if the default
#  (integer-based calculation) is not suitable
Llin.rep.conf.sort = confPlot(Llin.rep.df[ ,-1], legName="(a) Llin",
    xLim = c(-0.25, 0.25)) #, colourCode=FALSE)

LT.rep.conf.sort = confPlot(LT.rep.df[ ,-1], legName="(b) LT", yLab="", yLabels=FALSE)

LTplus1.rep.conf.sort = confPlot(LTplus1.rep.df[ ,-1], legName="(c) LTplus1")

xLimCom = c(-2.65, -1.5)     # common width of axes for the remainder

# range(LBmiz.rep.conf) =  -1.5894281 -0.5268428. Was -1.6, -0.4
LBmiz.rep.conf.sort = confPlot(LBmiz.rep.df[ ,-1] - 1, legName="(d) LBmiz",
    xLim = xLimCom, yLab="", yLabels=FALSE)

# range(LBbiom.rep.conf) = -0.6106510  0.4370617
LBbiom.rep.conf.sort = confPlot(LBbiom.rep.df[ ,-1] - 2, legName="(e) LBbiom",
      xLim = xLimCom)

# range(LBNbiom.rep.conf) = -1.6106510 -0.5629383
LBNbiom.rep.conf.sort = confPlot(LBNbiom.rep.df[ ,-1] - 1, legName="(f) LBNbiom",
     yLab="", yLabels=FALSE, xLim = xLimCom)

# range(LCD.rep.conf) = -1.1744680 -0.8713301   was c(-1.2, -0.8)
LCD.rep.conf.sort = confPlot(LCD.rep.df[ ,-1] - 1, legName="(g) LCD",
    xLim = xLimCom, vertFirst = TRUE)

# c(-2.2, -1.8) - looks good, but better to have consistent
MLE.rep.conf.sort = confPlot(MLE.rep.df[ ,-1], legName="(h) MLE",
    xLim = xLimCom, yLab="", yLabels=FALSE)
mtext(expression(paste("Estimate of ", italic(b)), sep=""),
      side=1, outer=TRUE, line=-0.2, cex=0.8)
```

And Figures TODO for the MLE and MLEfix methods

```{r}
# Confidence intervals for MLE and MLEfix methods.
#postscript("fitting3confMLEfix.eps", height = 0.8*figheight,
#           width = 0.8*figwidth,
#           horizontal=FALSE,  paper="special")
par(omi = c(0.12, 0.05, 0.12, 0.0))      # outer margins in inches
par(mfrow=c(2,1)) #7,1))

oldmai = par("mai")    #   0.6732 0.5412 0.5412 0.2772  inches I think,
                       #    may be indpt of fig size
par(mai=c(0.5, 0.5, 0.1, 0.3))  # Affects all four figures if don't change agaiin
par(xaxs="i", yaxs="i")    # Have to define here for hist
par(mgp=c(2.0, 0.5, 0))    # puts axes labels closer I think
par(cex = 0.8)             # With no option all text comes out a bit small

# xrangeMLEfix = c(-2.225, -1.775)  # Range for these two figures
# breakshistMLEfix = seq(xrangeMLEfix[1], xrangeMLEfix[2], by=3/61)

MLE.rep.conf.sort = confPlot(MLE.rep.df[ ,-1], legName="(a) MLE",
    xLim = xLimCom, insetVal = c(-0.04, -0.03), insetVal2 = c(-0.04, 0.05))


MLEfix.rep.conf.sort = confPlot(MLEfix.rep.df[ ,-1], legName="(b) MLEfix",
    xLim = xLimCom, insetVal = c(-0.04, -0.03), insetVal2 = c(-0.04, 0.05))

mtext(expression(paste("Estimate of ", italic(b)), sep=""),
      side=1, outer=TRUE, line=-0.2, cex=0.8)

```


Range of widths of confidence intervals for MLE method is
`r range(MLE.rep.df$confMax - MLE.rep.df$confMax)`.

Range of widths of confidence intervals for MLEfix method is
`r range(MLEfix.rep.df$confMax - MLEfix.rep.df$confMax)`.

<!--
CHECK THIS:   load(file="fitting2.RData")    # or load in data for x
From readMeCode.txt:

The resulting figures (as .eps postscript files) are also included so that
 code can be independnetly re-run and the results easily compared with my
 original figures. I have included the .RData file for the results for
 the main simulation of 10,000 data sets (multiple/fitting3rep.r)
 and for the xmax=10000 simulation (multiple/xmax10000/fitting3rep10000.r)
 because the code can take a while to run and the two resulting .RData files
 are used to produce Figure 3. The remainaing .RData files are have
 not been included because they are generally large.

The main figures in the manuscript can be found in the following directories:

Figures 1 and 2 - code/single/
Figures 3 and 4 - code/multiple/
Figure 5 - code/MLEbin/
Figure 6 - code/recommend/

So to use the MLE method to analyse your own data and plot results as per
 our Figure 6, see code/recommend/ . If your data are binned then you will
 need some of code/MLEbin/ .

code/fitting2.r  (only r file there):

# fitting2.r - replacing the fitting methods with functions, that are all
#  in PLBfunctions.r. 5th June 2015.


dev.off()

save.image(file = "fitting2.RData")
-->

# Figure 5 - MLEbin

Come back to once done from binning paper as likely more functionalised.

# Figure 6 - recommend

TODO see t'other vignette.
