---
title: "MEE_reproduce_1"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEE_reproduce_1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

```{r setup}
library(sizeSpectra)
```

# Reproduce Figures 1, 2 and A.1 of MEE paper

This vignette reproduces all the figures and tables in TODO: ref paper. See the
paper for details. The aim of the vignette is to enable users to reproduce the
results so that they can see what functions are appropriate for their own data,
and perform further analyses as required.


## Figure 1 - standard histogram of a random sample from a bounded power-law distribution

Generate random sample from bounded power-law distribution, where the resulting
`x` is a vector of individual fish sizes (usually body masses).
```{r}
n = 1000                  # sample size
b.known = -2              # known fixed value of b
xmin.known = 1            # known fixed value of xmin
xmax.known = 1000         # known fixed value of xmax
set.seed(42)              # To get the same observations for each run of code.

x = rPLB(n,
         b = b.known,
         xmin = xmin.known,
         xmax = xmax.known)

head(x)
```

Make the standard histogram with a break in the y-axis (by first fitting using
the Llin method since that returns the required bin breaks, counts etc.).
```{r}
num.bins = 8    # Suggested number of bins for standard histogram and Llin
                #  method; Daan et al. used 8 bins.

# hAAA is the h(istrogram) results for method AAA.
# Llin method - plotting binned data on log-linear axes then fitting regression.
hLlin = Llin.method(x,
                    num.bins = num.bins)

gap.barplot.cust(hLlin$counts,
                 midpoints = hLlin$mids,
                 breakpoints = hLlin$breaks,
                 xlim = c(-10,max(hLlin$breaks)+10),
                 col = rep("grey", length(hLlin$counts)),
                 xaxs = "i",
                 yaxs = "i"
                 )
```

## Figure 2 -- results from using eight methods to fit size spectra

Calculate size spectra fits using all eight methods on the same data set.
```{r}
eight.results = eightMethodsMEE(x, num.bins = num.bins)
```

The following code plots Figure 2, where each panel corresponds to one of the
methods. Most of this code involves details of the plots to make good figures,
and so might somewhat tailored for this explicit data set. So it may need
tweaking for another dataset. (Could be made into a function with lots of
arguments).

```{r, fig.height = 8}

par(mfrow=c(4,2))
oldmai = par("mai")
par(mai=c(0.4, 0.5, 0.05, 0.3))  # Affects all figures if don't change again
mgpVals = c(1.6,0.5,0)           # mgp values, default 2.0, 0.5, 0

# Notation:
# hAAA - h(istrogram) for method AAA.

# Llin method

hLlin.list = eight.results$hLlin.list

plot(hLlin.list$mids,
     hLlin.list$log.counts,
     xlim=c(0, max(hLlin.list$breaks)),
     xlab=expression(paste("Bin midpoints for data ", italic(x))),
     ylab = "Log (count)", mgp=mgpVals)

lm.line(hLlin.list$mids, hLlin.list$lm)
inset = c(0, -0.04)     # inset distance of legend

legJust(c("(a) Llin",
          paste("slope=",
                signif(hLlin.list$slope, 3),
                sep="")),
        inset=inset)

# LT method - plotting binned data on log-log axes then fitting regression,
#  as done by Boldt et al. 2005, natural log of counts plotted against natural
#  log of size-class midpoints.

hLT.list = eight.results$hLT.list

plot(hLT.list$log.mids,
     hLT.list$log.counts,
     xlab=expression(paste("Log (bin midpoints for data ", italic(x), ")")),
     ylab = "Log (count)",
     mgp=mgpVals)

lm.line(hLT.list$log.mids, hLT.list$lm)

legJust(c("(b) LT",
          paste("slope=", signif(hLT.list$slope, 3), sep=""),
          paste("b=", signif(hLT.list$slope, 3), sep="")),
        inset=inset)

# LTplus1 method - plotting linearly binned data on log-log axes then fitting
#  regression of log10(counts+1) vs log10(midpoint of bins), as done by
#  Dulvy et al. (2004).

hLTplus1.list = eight.results$hLTplus1.list

plot(hLTplus1.list$log10.mids,
     hLTplus1.list$log10.counts,
     xlab=expression(paste("Log10 (bin midpoints for data ", italic(x), ")")),
     ylab = "Log10 (count+1)",
     mgp=mgpVals,
     yaxt="n")

if(min(hLTplus1.list$log10.counts) < par("usr")[3]
      | max(hLTplus1.list$log10.counts) > par("usr")[4])
   { stop("fix ylim for LTplus1 method")}

axis(2,
     at = 0:3,
     mgp=mgpVals)
axis(2,
     at = c(0.5, 1.5, 2.5),
     mgp=mgpVals,
     tcl=-0.2,
     labels=rep("", 3))

lm.line(hLTplus1.list$log10.mids, hLTplus1.list$lm)
legJust(c("(c) LTplus1",
          paste("slope=", signif(hLTplus1.list$slope, 3), sep=""),
          paste("b=", signif(hLTplus1.list$slope, 3), sep="")),
        inset=inset)

# LBmiz method - binning data using log10 bins, plotting results on natural
#  log axes (as in mizer). Mizer does abundance size spectrum or biomass
#  size spectrum - the latter multiplies abundance by the min of each bin.

hLBmiz.list = eight.results$hLBmiz.list

plot( hLBmiz.list$log.min.of.bins, hLBmiz.list$log.counts,
     xlab=expression(paste("Log (minima of bins for data ", italic(x), ")")),
     ylab = "Log (count)", mgp=mgpVals)

lm.line(hLBmiz.list$log.min.of.bins, hLBmiz.list$lm)
legJust(c("(d) LBmiz",
          paste("slope=", signif(hLBmiz.list$slope, 3), sep=""),
          paste("b=", signif(hLBmiz.list$slope - 1, 3), sep="")),
        inset=inset)

# LBbiom method - binning data using log2 bins, calculating biomass (not counts)
#  in each bin, plotting log10(biomass in bin) vs log10(midpoint of bin)
#  as done by Jennings et al. (2007), who used bins defined by a log2 scale.

hLBNbiom.list = eight.results$hLBNbiom.list  # This does LBbiom and LBNbiom methods.

plot(hLBNbiom.list[["binVals"]]$log10binMid,
     hLBNbiom.list[["binVals"]]$log10totalBiom,
     xlab=expression(paste("Log10 (bin midpoints for data ", italic(x), ")")),
     ylab = "Log10 (biomass)",
     mgp=mgpVals,
     xlim=c(0, 2.7),
     ylim=c(2.5, 3.0),
     yaxt="n")

if(min(hLBNbiom.list[["binVals"]]$log10binMid) < par("usr")[1]
    | max(hLBNbiom.list[["binVals"]]$log10binMid) > par("usr")[2])
   { stop("fix xlim for LBbiom method")}

if(min(hLBNbiom.list[["binVals"]]$log10totalBiom) < par("usr")[3]
    | max(hLBNbiom.list[["binVals"]]$log10totalBiom) > par("usr")[4])
   { stop("fix ylim for LBbiom method")}

axis(2, at = seq(2.5, 3, 0.25), mgp=mgpVals)
axis(2, at = seq(2.5, 3, 0.05), mgp=mgpVals, tcl=-0.2, labels=rep("", 11))

lm.line(hLBNbiom.list[["binVals"]]$log10binMid, hLBNbiom.list[["unNorm.lm"]])

legJust(c("(e) LBbiom",
          paste("slope=", signif(hLBNbiom.list[["unNorm.slope"]], 1), sep=""),
          paste("b=", signif(hLBNbiom.list[["unNorm.slope"]] - 2, 3), sep="")),
        inset=inset)

# LBNbiom method - on biomass, not counts, as per Julia Blanchard's 2005 paper.
#  log2 bins of bodymass, sum the total biomass in each bin, normalise
#  biomasses by binwidths, fit regression to log10(normalised biomass) v
#  log10(midpoint of bin).

plot(hLBNbiom.list[["binVals"]]$log10binMid,
     hLBNbiom.list[["binVals"]]$log10totalBiomNorm,
     xlab=expression(paste("Log10 (bin midpoints for data ", italic(x), ")")),
     ylab = "Log10 (normalised biomass)",
     mgp=mgpVals,
     xlim=c(0,2.7),
     ylim=c(0,3),
     yaxt="n")

if(min(hLBNbiom.list[["binVals"]]$log10binMid) < par("usr")[1]
    | max(hLBNbiom.list[["binVals"]]$log10binMid) > par("usr")[2])
{ stop("fix xlim for LBNbiom method")}

if(min(hLBNbiom.list[["binVals"]]$log10totalBiomNorm) < par("usr")[3]
    | max(hLBNbiom.list[["binVals"]]$log10totalBiomNorm) > par("usr")[4])
   { stop("fix ylim for LBNbiom method")}

axis(2, at = 0:3, mgp=mgpVals)
axis(2, at = c(0.5, 1.5, 2.5), mgp=mgpVals, tcl=-0.2, labels=rep("", 3))

lm.line(hLBNbiom.list[["binVals"]]$log10binMid, hLBNbiom.list[["norm.lm"]])

legJust(c("(f) LBNbiom",
          paste("slope=", signif(hLBNbiom.list[["norm.slope"]], 3), sep=""),
          paste("b=", signif(hLBNbiom.list[["norm.slope"]] - 1, 3), sep="")),
        inset=inset)

# Cumulative Distribution, LCD method

hLCD.list = eight.results$hLCD.list

plot(hLCD.list$logSorted,
     hLCD.list$logProp,
     main="",
     xlab=expression(paste("Log ", italic(x))),
     ylab=expression( paste("Log (prop. of ", values >= italic(x), ")")),
     mgp=mgpVals)
     #xlim=c(0.8, 1000), xaxs="i", ylim=c(0.0008, 1), yaxs="i",

lm.line(hLCD.list$logSorted, hLCD.list$lm, col="red")

legJust(c("(g) LCD",
          paste("slope=", signif(hLCD.list$slope, 3), sep=""),
          paste("b=", signif(hLCD.list$slope - 1, 3), sep="")),
        inset=inset)
# MLE (maximum likelihood method) calculations.

hMLE.list = eight.results$hMLE.list

xmin = min(x)    # To avoid keep calculating
xmax = max(x)

# To plot rank/frequency style plot:
plot(sort(x, decreasing=TRUE), 1:length(x), log="xy",
     xlab=expression(paste("Values, ", italic(x))),
     ylab=expression( paste("Number of ", values >= x), sep=""),
     mgp=mgpVals, xlim = c(xmin, xmax), ylim = c(1, n), axes=FALSE)
xLim = 10^par("usr")[1:2]
yLim = 10^par("usr")[3:4]

logTicks(xLim, yLim, xLabelSmall = c(5, 50, 500))   # Tick marks.

x.PLB = seq(min(x), max(x), length=1000)     # x values to plot PLB
y.PLB = (1 - pPLB(x = x.PLB,
                  b = hMLE.list$b,
                  xmin = min(x.PLB),
                  xmax = max(x.PLB))) * length(x)
lines(x.PLB, y.PLB, col="red")     #, lty=5)
legJust(c("(h) MLE",
          paste("b=", signif(hMLE.list$b, 3), sep="")),
        inset=inset,
        logxy=TRUE)
```

## Figure A.1

Same as (h) above (Figure 2(h) in paper), but on linear-log axes, to help explain in the
why the red curve on log-log axes in (h) above does not pass through the maximum
data point.

```{r, fig.height = 5.4}
par(mai=c(0.8, 0.8, 0.2, 0.3))
mgpVal = c(2, 0.5, 0)
plot(sort(x,
          decreasing=TRUE),
     1:length(x),
     log="x",
     xlab=expression(paste("Values, ", italic(x))),
     ylab=expression( paste("Number of ", values >= x), sep=""),
     xlim = c(xmin, xmax), ylim = c(0, n), axes=FALSE,  mgp=mgpVal)
xLim = 10^par("usr")[1:2]
yLim = NULL

logTicks(xLim,
         yLim = NULL,
         xLabelSmall = c(5, 50, 500),
         mgp = mgpVal)

yBig = c(0, 500, 1000)
# Big labelled:
axis(2,
     at = yBig,
     labels = yBig,
     mgp = mgpVal)
# Small unlabelled:
axis(2,
     seq(yBig[1], yBig[length(yBig)], by = 100),
     labels = rep("", 11),
     tcl = -0.2,
     mgp = mgpVal)

x.PLB = seq(min(x), max(x), length=1000)     # x values to plot PLB
y.PLB = (1 - pPLB(x = x.PLB,
                  b = hMLE.list$b,
                  xmin = min(x.PLB),
                  xmax = max(x.PLB))) * length(x)
lines(x.PLB, y.PLB, col="red") #, lty=5)
```


<!--
CHECK THIS:   load(file="fitting2.RData")    # or load in data for x
From readMeCode.txt:

The resulting figures (as .eps postscript files) are also included so that
 code can be independnetly re-run and the results easily compared with my
 original figures. I have included the .RData file for the results for
 the main simulation of 10,000 data sets (multiple/fitting3rep.r)
 and for the xmax=10000 simulation (multiple/xmax10000/fitting3rep10000.r)
 because the code can take a while to run and the two resulting .RData files
 are used to produce Figure 3. The remainaing .RData files are have
 not been included because they are generally large.

The main figures in the manuscript can be found in the following directories:

Figures 1 and 2 - code/single/
Figures 3 and 4 - code/multiple/
Figure 5 - code/MLEbin/
Figure 6 - code/recommend/

So to use the MLE method to analyse your own data and plot results as per
 our Figure 6, see code/recommend/ . If your data are binned then you will
 need some of code/MLEbin/ .

code/fitting2.r  (only r file there):

# fitting2.r - replacing the fitting methods with functions, that are all
#  in PLBfunctions.r. 5th June 2015.


dev.off()

save.image(file = "fitting2.RData")
-->
