---
title: "MEPS_IBTS_MLEbins"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEPS_IBTS_MLEbins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 7
)
```

## Analyses of IBTS data using the MLEbins method

This vignette analyses the IBTS data using the MLEbins method that explicitly accounts for
the species-specific body-mass bins.

Creates Figure 6 (and related Figures S.1, S.2 and S.3) showing
species-specific body mass bins resulting from the length bins, Figure 8
(comparison of MLE and MLEbins values of b through time) and MLEbins row of
Table S.1.

```{r}
library(sizeSpectra)
library(tibble)  # Else prints all of a tibble
data = IBTS_data
data
```

## Determining which rows are 0.5 cm bins

`LngtClass` for all species is the minimum value of a 1-cm-width bin, except for
herring (*Clupea harengus*) and sprat (*Sprattus sprattus*) for which lengths
are rounded down to 0.5 cm values (so the bins are 0.5-cm wide). The `SpecCode`
values for these are:

```{r speciesNames}
herringCode = dplyr::filter(specCodeNames, species == "Clupea harengus")$speccode
herringCode
spratCode = dplyr::filter(specCodeNames, species == "Sprattus sprattus")$speccode
spratCode
specCode05 = c(herringCode, spratCode)      # species codes with 0.5cm length bins
```

Verified earlier that only these two species have 0.5 cm values for `LngtClass`.

## Append the max of the bin breaks for each row

So `LngtClass` is the minimum of each length bin. Need to work out the maximum
of each length bin `LengthMax`, and then use the species-specific length-weight
relationships to give the min (`wmin`) and max (`wmax`) of each body-mass bin.
So create `dataBin` table dataframe that has `LengthMax`, `wmin` and `wmax` as
extra columns for each row:

```{r dataBin}
dataBin = dplyr::mutate(data,
                        LngtMax = LngtClass + 1)
aa = which(dataBin$SpecCode %in% specCode05)           # row numbers for herring, sprat
dataBin[aa, "LngtMax"] = dataBin[aa, "LngtMax"] - 0.5  # subtract 0.5 cm to
                                                       # give 0.5-cm wide bins
unique(dataBin$LngtMax - dataBin$LngtClass)            # correctly just has 0.5 and 1
unique( dplyr::filter(dataBin, LngtMax - LngtClass == 0.5)$SpecCode)  # just herring,sprat

dataBin = dplyr::mutate(dataBin, wmax = LWa * LngtMax^LWb)  # calculate max body mass
                                                            # for each bin (min
                                                            # is currently bodyMass)
dataBin = dplyr::rename(dataBin, LngtMin = LngtClass)       # For consistency
dataBin = dplyr::rename(dataBin, wmin = bodyMass)

dataBin = dataBin[ , c("Year", "SpecCode", "LngtMin", "LngtMax",
                       "LWa", "LWb", "wmin", "wmax", "Number")]     # Reorder columns

range(dplyr::mutate(dataBin,
                    wminCheck = LWa * LngtMin^LWb)$wminCheck - dataBin$wmin)
                                              # Verifying that wmin is correct
                                              # (was calculated independently)
length(unique(dataBin$SpecCode))
```

This is the code to then save `dataBin` as a data set in the package, but is not
run here:
```{r dataBinsave, eval=FALSE}
usethis::use_data(dataBin, overwrite = TRUE)
```

## Plot the resulting body mass bins

So there are `r length(unique(dataBin$SpecCode))` uniques species. Now going to
plot the resulting body mass bins for each species, with 45 on each figure. This
gives Figures 6, S.1, S.2 and S.3. This function wrangles the data, calculates
some useful values and plots all four figures:

```{r fig.width=7.5, fig.height=6}
res <- species_bins_plots()
```

Those four figures show how the length bins for each species get converted to
body mass bins. The conversions are different for each species because of the
different values of the length-weight coefficients. Even with 1-cm length bins
(and 0.5-cm for herring and sprat) the resulting body-mass bins can span a large
range. See paper for further details.

Two species are highlighted: *Triglops murrayi* is Moustache Sculpin (code 127205). *Lumpenus lampretaeformis* is Snakeblenny (code 154675). Data for these are:
```{r highlight}
dataHighlight = dplyr::filter(data,
                              SpecCode %in% c(127205, 154675))
dataHighlightSumm = dplyr::summarise(dplyr::group_by(dataHighlight,
                                                     SpecCode),
                                     minLngt = min(LngtClass),
                                     maxLngt = max(LngtClass),
                                     LWa = unique(LWa),
                                     LWb = unique(LWb))
dataHighlightSumm
```

The widest resulting body-mass bin occurs for Atlantic Cod (`r res$specNameMaxWidth`), which is
the rightmost species in the final figure above figure. The widest bin
has a width of `r round(res$maxWidth, dig=0)` g.


```{r echo=FALSE, eval=FALSE}
# Plot of wmax for each species. I think it's a metric used somewhere.
plot(1:(dim(dataBinSpecWmax)[1]),
     dataBinSpecWmax$maxWmax,
     log="y",
     xlab="Species index",
     ylab="Wmax for each species")

# Plot of maximum overall body size (max of max(wmax)) each year
maxWmaxByYear = dplyr::summarise(dplyr::group_by(dataBin, Year),
                                 maxWmax = max(wmax))
smallTck=0.01
yLimWmaxByYear = range(pretty(c(0, max(maxWmaxByYear$maxWmax))))
plot(maxWmaxByYear$Year,
     maxWmaxByYear$maxWmax,
     xlab="Year",
     ylab="Wmax for each year",
     ylim=yLimWmaxByYear)
xTicksSmall = maxWmaxByYear$Year
axis(1, at = xTicksSmall, labels = rep("", length(xTicksSmall)), tck=-smallTck)
yTicksSmall = seq(yLimWmaxByYear[1], yLimWmaxByYear[2], by=2000)
axis(2, at = yTicksSmall, labels = rep("", length(yTicksSmall)), tck=-smallTck)
# points(2011, dplyr::filter(maxWmaxByYear, Year == 2011)$maxWmax, col="red", pch=19)
```

## Likelihood calculations using new MLEbins method

Now use the MLEbins method to fit each year of data in turn.

```{r MLEbins}
fullYears = sort(unique(dataBin$Year))
# Do a loop for each year, saving all the results in MLEbins.nSeaFung.new
for(iii in 1:length(fullYears))
  {
    dataBinForLike = dplyr::filter(dataBin,
                                   Year == fullYears[iii])
    dataBinForLike = dplyr::select(dataBinForLike,
                                   SpecCode,
                                   wmin,
                                   wmax,
                                   Number)
    n = sum(dataBinForLike$Number)
    xmin = min(dataBinForLike$wmin)
    xmax = max(dataBinForLike$wmax)

    MLEbins.nSeaFung.oneyear.new  = calcLike(negLL.fn = negLL.PLB.bins.species,
                                             p = -1.9,
                                             suppress.warnings = TRUE,
                                             dataBinForLike = dataBinForLike,
                                             n = n,
                                             xmin = xmin,
                                             xmax = xmax)

    if(iii == 1)
    {
      MLEbins.nSeaFung.new = data.frame(Year = fullYears[iii],
                                        xmin=xmin,
                                        xmax=xmax,
                                        n=n,
                                        b=MLEbins.nSeaFung.oneyear.new$MLE,
                                        confMin=MLEbins.nSeaFung.oneyear.new$conf[1],
                                        confMax=MLEbins.nSeaFung.oneyear.new$conf[2])
    } else {
      MLEbins.nSeaFung.new = rbind(MLEbins.nSeaFung.new,
                                   c(fullYears[iii],
                                     xmin,
                                     xmax,
                                     n,
                                     MLEbins.nSeaFung.oneyear.new$MLE,
                                     MLEbins.nSeaFung.oneyear.new$conf[1],
                                     MLEbins.nSeaFung.oneyear.new$conf[2]))
   }
}

# Need the standard error for weighted linear regression,
#  see eightMethods.count() for details:
MLEbins.nSeaFung.new = dplyr::tbl_df(MLEbins.nSeaFung.new)
MLEbins.nSeaFung.new = dplyr::mutate(MLEbins.nSeaFung.new,
                                     stdErr = (abs(confMin-b) +
                                               abs(confMax-b))/(2*1.96) )
MLEbins.nSeaFung.new
```

Now to plot the results and obtain the regression fit:
```{r timeseries, fig.width=7.5, fig.height=6}
res = timeSerPlot(MLEbins.nSeaFung.new,
                  legName = "(a) MLEbins",
                  yLim = c(-2.2, -0.9),
                  xLab = "Year",
                  method = "MLEbins",
                  legPos = "bottomleft",
                  weightReg = TRUE,
                  xTicksSmallInc = 1,
                  yTicksSmallInc = 0.05)
```

The statistics for the regression fit, the final row in Table S.1 are:
```{r}
trendResultsMLEbinsNew = dplyr::tbl_df(res)
knitr::kable(dplyr::select(trendResultsMLEbinsNew, Method, Low, Trend, High, p, Rsquared),
             digits=c(NA, 4, 4, 4, 2, 2))
```

And use the results to plot Figure 8, comparing results from the original MLE
method with those from the MLEbins method.
```{r fig.width=7.5, fig.height=6.3}
fullResults.MLEbins = MLEbins.nSeaFung.new  # Should really have just used
                                        # MLEbins..; happened to include nSeaFung early on
trend.MLEbins.new = dplyr::filter(trendResultsMLEbinsNew,
                                  Method == "MLEbins")
fullResults.MLE = dplyr::filter(fullResults, Method == "MLE")

bYears = fullResults.MLE$Year
MLE.col = "blue"
MLEbins.col = "red"
res.MLE = timeSerPlot(fullResults.MLE,
                      legName = "",
                      xLim = range(bYears),
                      yLim = c(-1.82, -1.35),
                      xLab = "Year",
                      method = "MLE",
                      legPos = "bottomleft",
                      weightReg = TRUE,
                      bCol = MLE.col,
                      pchVal = 19,
                      regPlot = FALSE,
                      regColNotSig = "lightblue",
                      regColSig = "darkblue",
                      xTicksSmallInc = 1,
                      yTicksSmallInc = 0.02,
                      legExtra = c("MLEbins", "MLE"),
                      legExtraCol = c(MLEbins.col, MLE.col),
                      legExtraPos = "topleft",
                      xJitter = -0.03)       # MLEbins on top as values are higher in figure

res.MLEbins.new = timeSerPlot(fullResults.MLEbins,
                              legName = "",
                              method = "MLEbins",
                              weightReg = TRUE,
                              newPlot = FALSE,
                              bCol = MLEbins.col,
                              confCol = MLEbins.col,
                              pchVal = 19,
                              regPlot = FALSE,
                              regColNotSig = "pink",
                              regColSig = "darkred",
                              xJitter = 0.03)
```

Finally, some further calculations to compare trends. This gives the ratio of
the estimates of *b* between the two methods:
```{r bRatio, fig.width=7.5, fig.height=6.3}
bRatio = fullResults.MLE$b / fullResults.MLEbins$b
xLim = range(bYears)
yLim = c(1, max(bRatio))
smallTck = 0.01

lm.bRatio = lm(bRatio ~ bYears)
yearInc = seq(xLim[1], xLim[2], 0.1)
p.conf.bRatio = predict(lm.bRatio,
                        newdata=data.frame(Year=bYears),
                        interval="confidence")
pVal.bRatio = summary(lm.bRatio)$coeff["bYears", 4]

plot( bYears,
     bRatio,
     xlab="Year",
     pch=19,
     ylab= expression(paste("(",
                            italic(b),
                            " from MLE)/(",
                            italic(b),
                            " from MLEbins)"),
                      sep=""),
     ylim = yLim)
xTicksSmall = seq(xLim[1], xLim[-1], by=1)
axis(1,
     at = xTicksSmall,
     labels = rep("", length(xTicksSmall)),
     tck=-smallTck)
yTicksSmall = seq(yLim[1], yLim[-1], by=0.05)
axis(2,
     at = yTicksSmall,
     labels = rep("", length(yTicksSmall)),
     tck=-smallTck)

if (pVal.bRatio > 0.05) regCol = "darkgrey" else regCol = "red"
lm.line(xLim, lm.bRatio, col=regCol)
matlines(bYears,
         p.conf.bRatio[ ,c("lwr", "upr")],
         col=regCol,
         lty=2)
confVals = confint(lm.bRatio, "bYears", level=0.95)
```
Above figure shows the annual ratio of the estimate of *b* using the original
MLE method to the estimate of *b* using the MLEbins method. The linear
regression slope is `r round(lm.bRatio$coeff[2],dig=5)` with 95\% confidence
interval `r round(confVals[1],dig=5)` to `r round(confVals[2],dig=5)`, *p*=`r
round(pVal.bRatio, dig=5)`, *R*^2=`r round(summary(lm.bRatio)$r.squared, dig=3)`.


And this shows the difference of the estimates of *b* between the two methods.
```{r bDiff, fig.width=7.5, fig.height=6.3}
bDiff = fullResults.MLEbins$b - fullResults.MLE$b
yLim = range(pretty(bDiff))

lm.bDiff = lm(bDiff ~ bYears)
p.conf.bDiff = predict(lm.bDiff,
                       newdata=data.frame(Year=bYears),
                       interval="confidence")
pVal.bDiff = summary(lm.bDiff)$coeff["bYears",4]

plot(bYears,
     bDiff,
     xlab="Year",
     pch=19,
     ylab= expression(paste("(",
                            italic(b),
                            " from MLEbins) - (",
                            italic(b),
                            " from MLE)"),
                      sep=""),
     ylim = yLim)
xTicksSmall = seq(xLim[1], xLim[-1], by=1)
axis(1,
     at = xTicksSmall,
     labels = rep("", length(xTicksSmall)),
     tck=-smallTck)
yTicksSmall = seq(yLim[1], yLim[-1], by=0.01)
axis(2,
     at = yTicksSmall,
     labels = rep("", length(yTicksSmall)),
     tck=-smallTck)

if (pVal.bDiff > 0.05) regCol = "darkgrey" else regCol = "red"
lm.line(xLim, lm.bDiff, col=regCol)
matlines(bYears,
         p.conf.bDiff[ ,c("lwr", "upr")],
         col = regCol,
         lty=2)
confVals = confint(lm.bDiff, "bYears", level=0.95)
```
The above figure shows the annual difference of the estimate of *b* using
MLEbins to the estimate of *b* using the original MLE method. The linear
regression slope is `r round(lm.bDiff$coeff[2],dig=5)` with 95\% confidence
interval `r round(confVals[1],dig=5)` to `r round(confVals[2],dig=5)`, *p*=`r
round(pVal.bDiff, dig=5)`, *R*^2=`r round(summary(lm.bDiff)$r.squared, dig=3)`.

TODO: THINK this can all go, S.2 done belowFinally, the actual results for each method, with the MLEbins table being Table
S.2:
TODO: use kable, first one might work, do the second one and edit to give Table
S.2 -
```{r, eval=FALSE}
#knitr::kable(dplyr::select(fullResults.MLE, -Method),
#             caption="Full results for original MLE method.",
#             lab="tab:fullresults.MLE",
#            digits=c(0, 0, 3, 3, 3, 5))

#print(fullResults.MLEtab, table.placement="tp", caption.placement="top",
#    include.rownames = FALSE, sanitize.text.function=function(x){x})  # was !ht


#fullResults.MLEbinsTab = xtable(
#  # dplyr::select(fullResults.MLEbins, -n, -xmin, -xmax),
#  fullResults.MLEbins,
#  caption="Full results for MLEbins.",
#  lab="tab:fullresults.MLEbins", digits=c(0, 0, 2, 2, 1, 3, 3, 3, 5))

# print(fullResults.MLEbinsTab, table.placement="tp", caption.placement="top",
#    include.rownames = FALSE, sanitize.text.function=function(x){x})  # was !ht
```

This is the start of `nSeaMLEbins-recommend.Snw`, maybe do figures in separate
vignette: [TODO - remove that]

Need the constant *C* for each year, so calculate it here:
```{r}
MLEbins.res = MLEbins.nSeaFung.new
MLEbins.res = dplyr::mutate(MLEbins.res,
                            C = (b != -1 ) * (b+1) / ( xmax^(b+1) - xmin^(b+1) ) +
                                (b == -1) * 1 / ( log(xmax) - log(xmin) )
                           )
MLEbins.res = dplyr::select(MLEbins.res, -stdErr)
```

This is Table S.2, given the results for each year from the MLEbins method:
```{r}
knitr::kable(dplyr::select(MLEbins.res, Year, xmin, xmax, n, confMin, b,
                           confMax, C),
             digits=c(0, rep(2, 7)))
```

```{r save}
# TODO - work out what to save, just copy from data-raw I expect.
# save.image(file="nSeaMLEbins.RData")

# MLEbins.res and dataBin saved in data-raw/IBTS-MLEbins.R

```
TODO: And see what to save in data-raw/IBTS-MLEbins.R, copy those lines to end
of here. maybe
