---
title: "MEPS_IBTS_recommend"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEPS_IBTS_recommend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 7
)
```

# Recommended plotting approach from the MLEbins method

This vignette has some of the calculations used to explain the recommended
plotting approach, reproduces Figure S.4, Figure 7 [TODO check] and presents
Figures S.5 to S.34 as a movie [TODO: hopefully].

```{r}
library(sizeSpectra)
# library(tibble)  # Else prints all of a tibble
```



TODO: And see what to save in data-raw/??.R, copy those lines to end of here,
may not need to save anything new

Copying nSeaMLEbins-recommend.Snw, some already in other vignette so deleting
that out:


## Schematic diagram for two species

Picking two example species (coloured in red in Figure 6). Codes are 127205
(Moustache Sculpin) and 154675 (Snakeblenny).

```{r}
sculpin = 127205
snakeblenny = 154675
specForFig = c(snakeblenny, sculpin)

dataTwoSpec = dplyr::filter(dataBin, SpecCode %in% specForFig)
```

Combine the data across all years for illustrative purposes (to reduce the number
of empty bins in the illustration):
```{r combine}
dataComb = dplyr::summarise(dplyr::group_by(dataTwoSpec,
                                            SpecCode,
                                            wmin),
                            wmax = unique(wmax),
                            number = round(sum(Number), dig=4))
dataComb = dplyr::mutate(dataComb, j = min_rank(wmin))   # works since they're grouped
dataComb[nrow(dataComb), "j"] = 25                       # manually, because there's a gap
dataComb = dplyr::mutate(dataComb, wmid = (wmin + wmax)/2)  # midpoint for plotting
```

Now create Figures S.4, with semi-automated text for the caption:
```{r fig.width=6.5, fig.height=4}
target = 7                 # target bin (for species 2)    # TODO 7 to target in
                           # next two lines
target.wmin = dplyr::filter(dataComb,
                            SpecCode == specForFig[2],
                            j == 7)$wmin
target.wmax = dplyr::filter(dataComb,
                            SpecCode == specForFig[2],
                            j == 7)$wmax
# Smallest and then largest possible inclusions of being > target bin:
dataComb = dplyr::mutate(dataComb,
                         wmin.gt.tmax = (wmin >= target.wmax))
dataComb = dplyr::mutate(dataComb,
                         wmax.gt.tmin = (wmax > target.wmin))

xLim = c(15, 43)           # range to show, must be integers
dataComb = dplyr::filter(dataComb,
                         wmin > xLim[1])  # only show complete bins

numSpec = length(specForFig)
yLimMax = 1
yVals = c(0.3, 0.7)        # y values for horizontal mass bins

colSpec = c("red", "pink")
thick = 30                 # y thickness of bins
cex.sub = 0.7              # font size for subscripts

#postscript("binCountRangeSchematic.eps",
#           height = 4, width = 6.5,
#           horizontal=FALSE,  paper="special")
par(mgp=c(2.0, 0.5, 0))    # puts axes labels closer I think
par(lend="butt")           # To have butted line caps, need for thick lines.
par(mar = c(4, 2.5, 3, 2)) # outer margins

plot(0, 0, xlab="Body mass, g", ylab="", xlim=xLim,
     ylim=c(0, yLimMax), yaxt="n", type="n", xaxs="i") # set up axes.

axis(1, at = xLim[1]:xLim[2], labels = rep("", xLim[2]-xLim[1]+1), tck=-0.01)

mtext("Species Code", side=2, line=1.5, cex.lab=1)
axis(2, at = yVals, labels = c(1,2), las = 2,
        tck = -0.005, cex.axis=0.8)

abline(v = target.wmin, col = "grey")
abline(v = target.wmax, col = "grey")

for(ii in 1:length(specForFig))   # loop over species, plot bins for each
  {
    yVal = yVals[ii]         # where to have horizontal bars
    bins.wmin = dplyr::filter(dataComb, SpecCode == specForFig[ii])$wmin  # for this sp
    bins.wmax = dplyr::filter(dataComb, SpecCode == specForFig[ii])$wmax
    bins.wmid = dplyr::filter(dataComb, SpecCode == specForFig[ii])$wmid
    # bins for which wmin >= target.wmax
    bins.wmin.gt.tmax = dplyr::filter(dataComb, SpecCode == specForFig[ii])$wmin.gt.tmax
    bins.wmin.gt.tmax.yn = ifelse(bins.wmin.gt.tmax,"Y", "N")
    # bins for which wmax > target.wmin
    bins.wmax.gt.tmin = dplyr::filter(dataComb, SpecCode == specForFig[ii])$wmax.gt.tmin
    bins.wmax.gt.tmin.yn = ifelse(bins.wmax.gt.tmin,"Y", "N")

    segments(y0 = yVal,
             x0 = bins.wmin,
             x1 = bins.wmax,
             col=colSpec, lwd=thick)    # recycles col
    # Bin break labels (cannot do , with vector for labels, it seems):
    for(iiii in 1:length(bins.wmin))
      {
        text(x = bins.wmin[iiii],
             y = yVal,
             labels = bquote(paste(w[.(ii)*","*
                    .(dplyr::filter(dataComb, SpecCode == specForFig[ii])$j[iiii])])),
             pos = 1,
             offset = 0.03*thick,
             cex = cex.sub)
      }
    # Do final wmax manually:
    text(x = bins.wmax[length(bins.wmax)],
         y = yVal,
         labels = bquote(paste(w[.(ii)*","*
                     .(max(dplyr::filter(dataComb, SpecCode == specForFig[ii])$j)+1)])),
         pos = 1,
         offset = 0.03*thick,
         cex = cex.sub)
    # Put counts for each bin
    text(x = bins.wmid,
         y = yVal,
         labels = f(dplyr::filter(dataComb, SpecCode == specForFig[ii])$number, ii+1),
                  # ii+1 happens to give 3 d.p.s for species 2 and 2 for species 1,
                  #  as needed on figure
         pos = NULL,
         offset = 0.03*thick,
         cex = cex.sub*1.2)
    # Whether to include in counts
    eps = 0.23                  # Offset from wmid for Y's and N's
    text(x = bins.wmid - eps,
         y = yVal,
         labels = bins.wmin.gt.tmax.yn,
         pos = 3,
         offset = 0.03*thick,
         cex = cex.sub*1.2)
    text(x = bins.wmid + eps,
         y = yVal,
         labels = bins.wmax.gt.tmin.yn,
         pos = 3,
         offset = 0.03*thick,
         cex = cex.sub*1.2)
  }

# Need w_1,24 manually, assumes this is the only one with j=23
text(x = dplyr::filter(dataComb, j==23)$wmax,
     y = yVals[1],
     labels = bquote(paste(w[.(1)*","*24])),
     pos = 1,
     offset = 0.03*thick,
     cex = cex.sub)

# Calculations for the caption:
lowCount = sum(dataComb$wmin.gt.tmax * dataComb$number)
highCount = sum(dataComb$wmax.gt.tmin * dataComb$number)

```

Caption: Schematic diagram to explain how we calculate
  the range of counts of individuals that are larger than those in a given bin.
  Red and pink body-mass bins are those for Snakeblenny and Moustache Sculpin
  (labelled species 1 and 2, respectively) from Figure 5 in the paper.
  Bin breaks are denoted by $w_{s,j}$
  and the number inside each bin is the number observed per hour of trawling.
  For illustration data are combined across all years and only bins with
  minima $>15$ g are shown. The target bin has $s=2$ and $j=7$ and therefore
  has bin breaks $w_{2,7}$ and $w_{2,8}$ and is indicated by the vertical grey
  lines.
  The first letter in each pair (`NN', `NY', or `YY') denotes whether or not
  each bin is included in the low count $E_{2,7,1}$, i.e. its lower bound is
  $\geq$ the upper bound of the target bin. Similarly, the second letter denotes
  whether or not each bin is included in the high count $E_{2,7,2}$, i.e.
  its upper bound is $>$ the lower bound of the target bin. Summing the
  respective counts as per (TODO ref{Esj1}) and (TODO ref{Esj1}) gives
  $E_{2,7,1} =$`r f(lowCount,2)` and $E_{2,7,2} =$ `r f(highCount,2)`.

TODO:
Figure~ref{fig:binCountRangeSchematic} shows these calculations for
one `target' bin ($s=2, j=7$) and two
example species namely Moustache Sculpin and Snakeblenny (which are highlighted
as 127205 and 154675, respectively, in red in Figure~5). Remaining text already
copied into Supp Material.

## Results plotted for each year

First just get what's needed for the calculations for the recommended plot:
TODO - should probably functionalise this if possible, would make input to
figure plotting consistent, and be easier for others to use:
```{r recommended}
dataRecommend.isd = dplyr::select(dataBin,
                                  Year,
                                  wmin,
                                  wmax,
                                  Number)

xRange.global = c(min(dataRecommend.isd$wmin),
                  max(dataRecommend.isd$wmax))
data.year.list = list()                # to save results for each year
diff.ivec = vector()                   # to save i that have any cumSum !=
                                       # verify TODO
fullYears = sort(unique(dataBin$Year))
for(i in 1:length(fullYears))
  {
    data.year = dplyr::filter(dataRecommend.isd,
                              Year == fullYears[i])

    data.year = dplyr::arrange(data.year,
                               desc(wmin))
    sumNumber = sum(data.year$Number)

    # data.year = dplyr::mutate(data.year,
    #                          cumSum = cumsum(Number))
    # This is wrong when we have two species with the same
    #  length-weight coefficients in the same year, use countGTEwmin then change it to
    #                          cumSum in one go TODO since clearer
    # Have to do not with dplyr:
    wmin.vec = data.year$wmin
    wmax.vec = data.year$wmax
    num.vec  = data.year$Number

    countGTEwmin = rep(NA, length(num.vec)) # to do a manual count
    lowCount = countGTEwmin
    highCount = countGTEwmin

    for(iii in 1:length(countGTEwmin))
    {
      countGTEwmin[iii]    = sum( (wmin.vec >= wmin.vec[iii]) * num.vec)
      lowCount[iii]  = sum( (wmin.vec >= wmax.vec[iii]) * num.vec)
      highCount[iii] = sum( (wmax.vec >  wmin.vec[iii]) * num.vec)
    }
    data.year = cbind(data.year,
                      "countGTEwmin" = countGTEwmin,
                      "lowCount" = lowCount,
                      "highCount" = highCount)
    data.year = dplyr::tbl_df(data.year)

    data.year.list[[i]] = data.year
}

```

Here is the code to give an animation for Figures 7 and S.5-S.34 for the IBTS data. For each year
(given in top-right corner) the plot shows the individual size distribution and
MLEbins fit (red solid curve) with 95\% confidence intervals (red dashed curves, may be hard to see).
For each bin, the horizontal green line shows the range of body sizes, with value on the y-axis
corresponding to the total number of individuals in bins whose minima are $\geq$
the bin's minimum.  For each bin, the vertical span of the grey rectangle shows
the possible range of the number of individuals with body mass $\geq$ the body
mass of individuals in that bin (horizontal span is the same as for the green
lines). The text in (a) gives the year, the MLE for the size-spectrum exponent
$b$, and the sample size $n$.

TODO Commenting out for now since gifski R package is causing Travis to be
upset.
<!-- ```{r, animation.hook = 'gifski', interval = 1.5, fig.width = 5.36, -->
<!-- fig.height = 8} -->
<!-- # fig.width is 0.67 * fig.height (which is 8)

for(i in 1:length(fullYears))   # TODO could make a function
  {
  ISD_bin_plot(data.year = data.year.list[[i]],
               b.MLE = dplyr::filter(MLEbins.res, Year == fullYears[i])$b,
               b.confMin = dplyr::filter(MLEbins.res, Year ==
                                                    fullYears[i])$confMin,
               b.confMax = dplyr::filter(MLEbins.res, Year ==
                                                    fullYears[i])$confMax,
               year = fullYears[i],
               xRange = xRange.global)
} -->
<!-- ```-->

```{r}
  # To save each year as it's own figure add something like this within the loop:
  # postscript(paste0(dirToSaveFigs, "/IBTS-ISD", fullYears[i], ".eps"),
  #             height = figheight, width = figwidth,
  #             horizontal=FALSE, paper="special")
  # ISD_bin_plot(...)     # include options like those above
  # dev.off()
  # See commented out latex code in vignette to give a loop to include each
  # figure, for example as done in our Supporting Information.
```

<!--
% Include this in latex preamble:
\newcommand\isdfig[2]{    % filename is #1, text is #2
  \begin{figure}[tp]
  \begin{center}
  \epsfbox{IBTS-ISDfigs/IBTS-ISD#1.eps}
  \end{center}
  \vspace{-5mm}
  \caption{#2}
  \label{fig:ISD-#1}
  \end{figure}
  \clearpage
}

% Include this after the figures have already been created in an R loop
\isdfig{\Sexpr{fullYears[1]}}{Include caption as above for year \Sexpr{fullYears[i]}}

\newcounter{loop}

\newcommand{\loopMax}{\Sexpr{max(fullYears)+1}}
\forloop{loop}{\Sexpr{fullYears[2]}}{\value{loop} < \loopMax}
        {\isdfig{\arabic{loop}}{Individual size distribution and MLEbins fit
            with 95\% confidence intervals for IBTS data in \arabic{loop}.
            Details as in Figure~\ref{fig:ISD-1986}.
        }}
-->
