---
title: "MLEbin_recommend"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MLEbin_recommend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 7
)
```

## Recommended fitting and plotting approach for data that are binned (with just one set of bin breaks)

```{r setup}
library(sizeSpectra)
# library(tibble)  # Else prints all of a tibble
```

This vignette is for data that are only available in standard binned form, with one
set of bin breaks (unlike the overlapping species-specific bins shown in MEPS
Fig. 7). For example, zooplankton data available in bins that progressively
double in bin width. Or, data that are only available in digitised form from earlier papers.

This vignette shows how to fit the PLB disribution (using the
MLEbin method) and plot it in a similar way to the MEPS Fig. 7. The plotting is
done using the new function `ISD_bin_plot_nonoverlapping()`, which is a wrapper
for the more complex (and slightly updated) `ISD_bin_plot()`.

Note: this is not (yet) part of any publication and so has not been peer-reviewed,
it's just exploring some potentially useful plotting approaches.

### Simulate and fit data

First, generate some data from a true PLB distribution and bin it (using bin widths
that double in size):
```{r generate}
set.seed(42)
x <- rPLB(n = 1000,
          b = -2,
          xmin = 1,
          xmax = 1000)

x.binned <- binData(x,
                    binWidth = "2k")  # 1, 2, 5 or "2k"

head(x.binned$indiv)     # Individual x values and which bin they are assigned to
x.binned$binVals         # Bins and the counts in each bin (with extra columns
                         #  that we won't need here).
```

Now fit the PLB distribution using the MLEbin method.
The fitting function takes the binned data as two vectors (the bin breaks
and the counts in each bin):
```{r likelihood}
num.bins <- nrow(x.binned$binVals)

# bin breaks are the minima plus the max of the final bin:
binBreaks <- c(dplyr::pull(x.binned$binVals, binMin),
               dplyr::pull(x.binned$binVals, binMax)[num.bins])

binCounts <- dplyr::pull(x.binned$binVals, binCount)

MLEbin.res <-  calcLike(negLL.fn = negLL.PLB.binned,
                        p = -1.5,
                        w = binBreaks,
                        d = binCounts,
                        J = length(binCounts),   # = num.bins
                        vecDiff = 1)             # increase this if hit a bound
```
The `NA/Inf replaced by maximum positive value` warnings are fine - the
likelihood function is blowing up in a very unlikely region of parameter
space.

### Plot the data and the fit like in MEPS Fig. 7

Now plot the data and results using:
```{r, plotvec, fig.width = 5.36, fig.height = 8}
# fig.width is 0.67 * fig.height (which is 8)
ISD_bin_plot_nonoverlapping(binBreaks = binBreaks,
                            binCounts = binCounts,
                            b.MLE = MLEbin.res$MLE,
                            b.confMin = MLEbin.res$conf[1],
                            b.confMax = MLEbin.res$conf[2])
```

The y-axis is (a) linear and (b) logarithmic. For each bin, the horizontal green
line shows the range of body masses of that bin, with its value
on the y-axis corresponding to the total number of individuals in bins whose
minima are $\geq$ the bin's minimum. By definition, this includes all individual
in that bin; for example, all $n = 1000$ individuals are $\geq$ the minimum of
the first (left-most) bin, so that bin is at $1000$ on the y-axis. The vertical
span of each grey rectangle shows the possible numbers of
individuals with body mass $\geq$ the
body mass of individuals in that bin (horizontal
span is the same as for the green lines). The maximum number of such individuals
is the same as the green line, and the minimum number (bottom of grey rectangle)
is the total number of individuals $\geq$ the bin's maximum. By definition, this is then the
green line for the next bin. This plotting method allows us to properly represent the
discrete binned data on a continuous scale. For example, for a body mass of 1.5
g we can see that the number (on the y-axis) of possible individual individuals with
body mass $\geq$ 1.5 g is between `r sum(binCounts) - binCounts[1]` and `r
sum(binCounts)`. This is true for all body masses in the first bin... HERE


The text
in (a) gives the MLE for the size-spectrum exponent
$b$, and the sample size $n$.

Both figures show that the PLB distribution is an excellent fit for these data
the red curves go through the top-left and bottom-right corners of the grey
boxes. Of course, the data are simulated from a PLB distribution and so should
be a good fit, but these figures show the benefit of the plotting approach
(which may not have been so obvious in MEPS Fig. 7 for data with a complex bin
structure).

### Plotting the data if stored as a tibble

Real data can either input to the plotting function in the above form - a vector `binBreaks`
of breakpoints and a vector `binCounts` of counts in each bin, or as a tibble.

If you prefer keeping your binned data as a tibble, then you can use the tible
in the call to the plotting function (though you do still need to input the data
as two vectors in the `calcLike` call above). This gives the same figures:
```{r, plottib, fig.width = 5.36, fig.height = 8}
ISD_bin_plot_nonoverlapping(binValsTibble = x.binned$binVals,
                            b.MLE = MLEbin.res$MLE,
                            b.confMin = MLEbin.res$conf[1],
                            b.confMax = MLEbin.res$conf[2],
                            yBig.inc = 500)
```

## LBN plot if data are body-masses

In MEE Fig. 6 we recommended, for body-mass data, showing the biomass size
spectrum, as people are used to seeing this type of plot. If body-mass data are only
available in binned form (as here), then this would be particularly
appropriate. The fit is calculated using likelihood, but then shown on the
biomass size spectrum.

Here, we introduce a modified version of MEE Fig 6(a). In
that figure the normalised biomass in each bin (the total biomass in a bin divided
by its bin width) was shown as a single point, as has been traditional in such
figures. However, this does not explicitly show the true width of the bin, and
ignores the uncertainty in the total biomass (that arises due to the binning).

Consider an example bin range from 15 to 20 g, with a count of 30
individuals. The possible range of the *total* biomass of those 30 individuals is
therefore 450 to 600 g, and the normalised biomass (dividing the total biomass
by the bin width) is 90 to 120.

More generally,
following the notation in MEE Section A.1.7, bin $j$ covers values of $x$ (body
masses) in the interval $[w_j, w_{j+1})$, and contains $d_j$ individuals. So the
total biomass in bin $j$ is in the range $[d_j w_j, d_j w_{j+1})$. The
normalised biomass in the bin (of width $w_{j+1} - w_j$), is therefore
in the range $d_j w_j /(w_{j+1} - w_j)$ to $d_j w_{j+1} / (w_{j+1} - w_j)$. We
express this as the vertical extent of the grey box for each bin in the figures below.

The horizontal range of each bin is simply shown as the horizontal extent of the
grey box for that bin; for bin $j$ this is simply $w_j$ to $w_{j+1}$, as in the
above figures.

We present three options, depending on whether axes are logged or not.

Neither axis logged:
```{r, plotLBN, fig.height = 5.7}
LBN_bin_plot(binValsTibble = x.binned$binVals,
             b.MLE = MLEbin.res$MLE,
             b.confMin = MLEbin.res$conf[1],
             b.confMax = MLEbin.res$conf[2],
             log.xy = "",
             plot.binned.fitted = FALSE)
```

This is clearly hard to see, hence motivates logging the x-axis:
```{r, plotLBN2, fig.height = 5.7}
LBN_bin_plot(binValsTibble = x.binned$binVals,
             b.MLE = MLEbin.res$MLE,
             b.confMin = MLEbin.res$conf[1],
             b.confMax = MLEbin.res$conf[2],
             leg.text = "(b)",
             log.xy = "x",
             plot.binned.fitted = FALSE)
```

This is easier to see and shows the good fit to the data, though it is harder to
see the high body-mass bins that have low normalised biomass in them. The grey
boxes show (horizontally) the range of each bin, and (vertically) the possible
range in normalised biomass for that bin, given the data. The red solid curve
fits through the boxes; dashed lines are the PLB distribution with $b$ set to
the limits of the 95% confidence interval (calculated using likelihood).

So log both axes (which is analogous to the traditional LBN plot, like in MEE
Fig. 6):
```{r, plotLBN3, fig.height = 5.7}
LBN_bin_plot(binValsTibble = x.binned$binVals,
             b.MLE = MLEbin.res$MLE,
             b.confMin = MLEbin.res$conf[1],
             b.confMax = MLEbin.res$conf[2],
             leg.text = "(c)",
             log.xy = "xy",
             plot.binned.fitted = FALSE)
```

The fit is really good for the numerous small individuals, less so for the
larger rare individuals. The log-log axes do emphasise the relatively poorer fit of the
larger body sizes, but there are few individuals here. The red lines are all
straight because of the log-log axes, but are calculated using likelihood, not
regression (which would be the LBNbiom method).

Since the data are only available in binned form, the straight red lines are,
upon reflection, not
actually the most intuitive things to plot. So next we estimate the expected
normalised biomass in each bin based on the MLE for $b$ (giving the horizontal
red bars) and the confidence limits for $b$ (giving the pink rectangles,
slightly shortened horizontally to still show the grey rectangles):
```{r, plotLBN4, fig.height = 5.7}
LBN_bin_plot(binValsTibble = x.binned$binVals,
             b.MLE = MLEbin.res$MLE,
             b.confMin = MLEbin.res$conf[1],
             b.confMax = MLEbin.res$conf[2],
             leg.text = "(c)",
             log.xy = "xy",
             plot.binned.fitted = TRUE)
```

This more clearly shows the uncertainty in expected normalised biomass is larger
at higher biomasses, because the bins are wider (I think). The pink and
grey boxes overlap, showing some consistency between them that is not obvious
from the earlier plots.

A useful thing will be to create movie where the fit is for incrementally higher
values of $x_{min}$


### Larger confidence interval curiosity

Just curious to see what the figure looks like with a confidence interval for
$b$ that's set to twice as wide as the actual one (fixing the interval here by
adding half the width to each side):
```{r, plotLBN22, fig.height = 5.7}
confWidth <- diff(MLEbin.res$conf)
confWidth
LBN_bin_plot(binValsTibble = x.binned$binVals,
             b.MLE = MLEbin.res$MLE,
             b.confMin = MLEbin.res$conf[1] - confWidth/2,
             b.confMax = MLEbin.res$conf[2] + confWidth/2)
```

But, now there's no real reason to label the axes with logged values, rather
than use the more intuitive actual values (still on a log scale).
