---
title: "MLEbin_recommend"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MLEbin_recommend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 7
)
```

## Recommended fitting and plotting approach for data that are binned

In the MEPS paper, we recommended using the MLEbin method for fitting data that
are only available in a binned form where the bins do not overlap each
other. This is the usual situation -- the more complex MLEbins method is for
when the bins overlap.

This vignette shows extracts the necessary code to fit and plot the results for
a general data set. This creates a simpler version of Fig. 7  -- simpler because
the green lines have to be at the top of the grey boxes because the bins do not
overlap CHECK???

TODO: maybe add some noise to data to add realism;

mention allows fitting of old data, digitised from plots

people have to get data in same form but I should keep it simple here
(binBreaks)


could try goodness-of-fit more naturally, from Ecology paper

```{r setup}
library(sizeSpectra)
# library(tibble)  # Else prints all of a tibble
```

Generate simulated data for this example; real data should be in the same
resulting format:
```{r generate}
# test <- MLEbin.simulate(num.reps = 1)  # that does way too much, copying the
# function here then simplifying down
n <- 1000      # sample size, but maybe calc later
b.known <- -2
xmin.known <- 1
xmax.known <- 1000

x <- rPLB(n,
          b = b.known,
          xmin = xmin.known,
          xmax = xmax.known)

binType = "2k"    # one of (1, 5, 10, "2k")  # can maybe be any number
#                           vecDiffVal = 0.5,
#                           cut.off = NA,
#                           full.mult = 1.5
#                          )
x.binned <- binData(x,
                    binWidth = binType)  # 1, 2, 5 or "2k"
head(x.binned$indiv)     # Individual x values and which bin they are assigned to
x.binned$binVals         # Bins and the counts in each bin

num.bins <- nrow(x.binned$binVals)

# bin breaks are the minima plus the max of the final bin:
binBreaks <- c(dplyr::pull(x.binned$binVals, binMin),
               dplyr::pull(x.binned$binVals, binMax)[num.bins])

# binBreaks <- c(binBreaks, maxOfMaxBin)             # Append endpoint of final
    # bin, maybe don't need
binCounts <- dplyr::pull(x.binned$binVals, binCount)

MLEbin.res <-  calcLike(negLL.PLB.binned,
                        p = -1.5,
                        w = binBreaks,
                        d = binCounts,
                        J = length(binCounts),   # = num.bins
                        vecDiff = 1)             # increase this if hit a bound

```

CHECK: Here is the code to give an animation for Figures 7 and S.5-S.34 for the IBTS data. For each year
(given in top-right corner) the plot shows the individual size distribution and
MLEbins fit (red solid curve) with 95\% confidence intervals (red dashed curves, may be hard to see).
For each bin, the horizontal green line shows the range of body sizes, with value on the y-axis
corresponding to the total number of individuals in bins whose minima are $\geq$
the bin's minimum.  For each bin, the vertical span of the grey rectangle shows
the possible range of the number of individuals with body mass $\geq$ the body
mass of individuals in that bin (horizontal span is the same as for the green
lines). The text in (a) gives the year, the MLE for the size-spectrum exponent
$b$, and the sample size $n$.

CHECK: Here is the code to build the movie, but it is commented out since it causes Travis
to fail -- Travis is the continuous integration service that automatically
checks the packages builds every time a change is committed to GitHub, giving the
little green symbol `passing` icon on the main page. The trick to showing the
movie is to run the code uncommented (you need to install the `gifski` package), right-click on the animation in the
html viewer and save it. I've done that and so am leaving this code commented,
and will just refer to the saved animation file.
```{r, fig.width = 5.36, fig.height = 8}
# fig.width is 0.67 * fig.height (which is 8)
#
ISD_bin_plot_nonoverlapping(binBreaks = binBreaks,
                            binCounts = binCounts,
                            b.MLE = MLEbin.res$MLE,
                            b.confMin = MLEbin.res$conf[1],
                            b.confMax = MLEbin.res$conf[2])

```
