---
title: "Vignettes_overview"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignettes_overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

```{r setup}
library(sizeSpectra)
```

# Overview of vignettes for sizeSpectra package

This R package contains functions for fitting size spectra to ecological
data. In particular, it contains functionalised code to reproduce all the
results in [1] and [2], and for users to apply the methods to their own data.

[1] **Testing and recommending methods for fitting size spectra to data** by
Andrew M. Edwards, James P. W. Robinson, Michael J. Plank, Julia K. Baum and
Julia L. Blanchard. ***Methods in Ecology and Evolution*** (2017,
8:57-67). Freely available at
<http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full>.

[2] **TODO** by Andrew M. Edwards, James P. W. Robinson, Julia L. Blanchard,
Julia K. Baum and Michael J. Plank. **TODO**

The vignettes explain how to use the functions in the package to reproduce the
results in both papers, to then enable users to see what functions are required
to analyse their own data.
The vignettes are descriptions of how to use the code to implement the
methods. The two papers should be consulted first to understand the methods (I
have tried to avoid repeating explanatory text from the papers in the vignettes). In the
vignettes, [1] is referred to as the 'MEE paper' and [2] as the 'MEPS paper'.

The vignettes are:

* [MEE_reproduce_1.html](MEE_reproduce_1.html): reproduces the figures in the
MEE paper that are based on the original single simulated data set, namely
Figures 1, 2, 6 and A.1.
