---
title: "Vignettes_overview"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignettes_overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

```{r setup}
library(sizeSpectra)
```

## Overview of vignettes for sizeSpectra package

This R package contains functions for fitting size spectra to ecological
data. In particular, it contains functionalised code to reproduce all the
results in [1] and [2], and for users to apply the methods to their own data.

[1] **Testing and recommending methods for fitting size spectra to data** by
Andrew M. Edwards, James P. W. Robinson, Michael J. Plank, Julia K. Baum and
Julia L. Blanchard. ***Methods in Ecology and Evolution*** (2017,
8:57-67). Freely available at
<http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12641/full>.

[2] **TODO** by Andrew M. Edwards, James P. W. Robinson, Julia L. Blanchard,
Julia K. Baum and Michael J. Plank. **TODO**

The vignettes explain how to use the functions in the package to reproduce the
results in both papers, to then enable users to see what functions are required
to analyse their own data.
The vignettes are descriptions of how to use the code to implement the
methods. The two papers should be consulted first to understand the methods (I
have tried to avoid repeating explanatory text from the papers in the vignettes). In the
vignettes, [1] is referred to as the 'MEE paper' and [2] as the 'MEPS paper'.

The vignettes are:

* [MEE_reproduce_1.html](MEE_reproduce_1.html): Simulate and fit a single data
  set using eight methods, and create recommended likelihood-based
  plot. Reproduces MEE Figures 1, 2, 6 and A.1.

* [MEE_reproduce_2.html](MEE_reproduce_2.html): Simulate 10,000
  data sets and fit them using eight methods, then plot histograms and confidence
  intervals of the estimates of the size-spectra exponent $b$. Reproduces MEE
  Figures 3 (TODO including gold?), 4 and Table 2. Also for the MLEfix method
  (Figures A.3 and A.4).

* [MEE_reproduce_3.html](MEE_reproduce_2.html): TODO not done yet, will
  be 3. Once 2 is functionalised, just repeat for different configurations.

* [MEPS_reproduce_1.html](MEPS_reproduce_1.html): Example length-weight
  relationships for two species and the effect of binned length data (MEPS
  Figure 2). And how consequent binned body-mass values get assigned to
  logarithimic size-class bins by some methods (Figure 3).

* [MEPS_reproduce_2.html](MEPS_reproduce_2.html): Main simulation results, from
  fitting maximum likelihood value (and confidence interval) of $b$ for each
  combination of 10,000 simulated data sets, four binning methods and two fitting
  methods (MLEmid or MLEbin). Reproduces MEPS Figures 4 and 5 and Table S.3. Then
  does the same analyses with $x$~min~$ = 16$ (Figures S.35 and S.36 and Table S.4),
  and with with $x$~min~$ = 1$ but only sampling above a cutoff value of 16
  (Figures S.37 and S.38 and Table S.5).

* [MEPS_IBTS_1.html](MEPS_IBTS_1.html): Pre-processing of International Bottom
  Trawl Survey (IBTS) data to extract the required elements. Gives MEPS Table 1.

* [MEPS_IBTS_2.html](MEPS_IBTS_2.html): Analyses of the IBTS data using the
  original eight methods. Gives MEPS Figure 1 and Table S.1.

* [MEPS_IBTS_MLEbins.html](MEPS_IBTS_MLEbins.html): Analyse the IBTS data using
  the MLEbins method. Gives Figure 6 (and related Figures S.1, S.2 and S.3)
  showing species-specific body mass bins resulting from the length bins, Figure
  8 (comparison of MLE and MLEbins values of b through time), MLEbins row of
  Table S.1, and Table S.2. Also some extra figures that helped understanding
  but didn't need to go into the paper.

* [MEPS_IBTS_recommend.html](MEPS_IBTS_recommend.html): Calculations and schematic diagram
  (Figure S.4, somewhat automated for any species) to help explain the
  recommended plotting approach. The recommended Figures 7, and S.5 to S.34 for
  each year of the IBTS data. TODO: either as movie or just a load of figures.
