---
title: "MEPS_IBTS_2"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEPS_IBTS_2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 6
)
```

## Analyses of IBTS data using the original eight methods

This vignette analyses the IBTS data using the original eight methods, without accounting for the bin structure.

```{r}
library(sizeSpectra)
data = IBTS_data
data
```

Each row of `data` is a unique combination of year, species code and length
class. The `Number` column is the number of observed individuals of that species
in that length class in that year, `bodyMass` is the body mass of such an
individual, and `LWa` and `LWb` are the length-weight coefficients, used to
calculate `bodyMass`. To verify that each row is a unique combination of year,
species code and length class:
```{r}
data
unique = dim(dplyr::summarise(dplyr::group_by(data,
                                              Year,
                                              SpecCode,
                                              LngtClass),
                              count=dplyr::n()))[1]
unique

if( unique != dim(data)[1]) stop("something wrong with 'data'")
```



See if the number of length classes or species seems to change over time:
```{r, dataSumm}
dataSumm = dplyr::summarise(dplyr::group_by(data, Year),
                            uniqLngtClass = length(unique(LngtClass)),
                            uniqSpec = length(unique(SpecCode)))
par(mfrow=c(2,1)) #7,1))

plot(dataSumm$Year, dataSumm$uniqLngtClass, xlab="Year",
     ylab="No. unique length classes", type="o",
     ylim=c(0, max(dataSumm$uniqLngtClass)))
plot(dataSumm$Year, dataSumm$uniqSpec, xlab="Year",
     ylab="No. unique species", type="o", ylim=c(0, max(dataSumm$uniqSpec)))
```

There do not look to be any serious issue with this (no drastic changes in, for
example, species identification through time).

The following code calculates, using `eightMethods.count()`, the slope or
exponent for each of the original eight methods for each year in turn, and saves a `.png` plot
 for each year that shows the fit of each method. It is
not run here because it takes a few hours (results are saved as the data object
`fullResults`).
```{r doEachYear, echo=TRUE, eval=FALSE}
fullYears = unique(data$Year)
fullResults = data.frame()
for(ii in fullYears){
  eightMethodsRes = eightMethods.count(data = data,
                                       oneYear = ii,
                                       figName = "nSeaFung")
  fullResults = rbind(fullResults, eightMethodsRes)
}
# usethis::use_data(fullResults, overwrite = TRUE)    # Run this to save the data.
```

Now to plot Figure 1, time series of each estimate of `b` for each method, and
fit regression:
```{r, fig.height = 7.5}
trendResults = timeSerPlot.eight()
```

The results of the regression analyses (Table S.1) are:
```{r results='asis'}
knitr::kable(dplyr::select(trendResults, -adjRsquared),
             digits=c(0, 4, 4, 4, 2, 2))
```
