---
title: "MEPS_reproduce_2"
author: "Andrew Edwards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MEPS_reproduce_2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.7,
  fig.height = 7
  )
```


Main simulation results for MEPS paper, from fitting maximum likelihood value (and confidence
interval) of $b$ for each  combination of 10,000 simulated data sets, four
binning methods and two fitting methods (MLEmid or MLEbin). Reproduces MEPS
Figures 4 and 5 and Table S.3. Then does the same analyses with $x$~min~$=16$
(Figures S.35 and S.36 and Table S.4), and with $x$~min~$=1$ but
only sampling above a cutoff value of 16 (Figures S.37 and S.38 and Table
S.5).

The simulated data take over an hour (on a fast laptop) to generate, bin and fit.
So results are saved within the package in the lists `MLEbin.MEPS.default`,
`MLEbin.MEPS.xmin16` and `MLEbin.MEPS.cutoff16`.
Each list contains the fitted maximum likelihood
value (and confidence interval) of $b$ for each combination of 10,000 simulated
data sets, four binning methods and two fitting methods (MLEmid or MLEbin).

```{r setup}
library(sizeSpectra)
```

## Reproducing main simulation results of MEPS paper -- Figures 4 and 5 and Table S.3

To create `MLEbin.MEPS.default` run the next line (which is not evaluated by
this vignette but is run once in `data-raw/simulate-data2.R` to save
the list as part of this package):
```{r, eval=FALSE}
MLEbin.MEPS.default <- MLEbin.simulate()
```

The list contains three items (see `?MLEbin.simulate`):
```{r}
summary(MLEbin.MEPS.default)
dim(MLEbin.MEPS.default$MLE.array)
```
with example MLEs of $b$:
```{r}
MLEbin.MEPS.default$MLE.array[ 1:5, , ]
```

Example confidence intervals are
```{r}
MLEbin.MEPS.default$MLEconf.array[1:5, "Linear 1", "MLEmid", ]
```
and the associated parameter values and binning types used for the simulations
are:
```{r}
MLEbin.MEPS.default$MLE.array.parameters
```


Figure 4 shows histograms of the estimates of $b$ for the MLEs for both methods and all four binning types:
```{r}
MLEmid.MLEbin.hist(MLEbin.MEPS.default)
```

Figure 5 gives the confidence intervals for the MLEs for both methods and all four binning types:
```{r}
# These two lines just give a different inset for panel (e):
insetMat = matrix(rep(c(-0.01, -0.04), 4),
                  ncol=2,
                  byrow=TRUE)
insetMat[3, 1] = 0.3

MLEmid.MLEbin.conf(MLEbin.MEPS.default,
                   insetMat = insetMat)
```

Make a dataframe of the results and show here, reproducing Table S.3:
```{r, echo=TRUE, results='asis'}
knitr::kable(MLEmid.MLEbin.table(MLEbin.MEPS.default))
```

## Setting $x$~min~$=16$ -- Figures S.35 and S.36 and Table S.4

Similarly, the simulation results obtained from setting $x$~min~$=16$ are saved in
the object `MLEbin.MEPS.xmin16`, obtained by running the next line (which is not evaluated by
this vignette but is run once in `data-raw/simulate-data2.R`):
```{r, eval=FALSE}
MLEbin.MEPS.xmin16 <- MLEbin.simulate(xmin.known = 16)
```

Figure S.35 is then
```{r}
MLEmid.MLEbin.hist(MLEbin.MEPS.xmin16)
```

Figure S.36 shows the confidence intervals:
```{r}
MLEmid.MLEbin.conf(MLEbin.MEPS.xmin16)
```

And Table S.4 is:
```{r, echo=TRUE, results='asis'}
knitr::kable(MLEmid.MLEbin.table(MLEbin.MEPS.xmin16))
```

## Setting $x$~min~$=1$ but only sampling data above the cutoff value of 16 -- Figures S.37 and S.38 and Table S.5)

Similarly, the simulation results obtained from setting $x$~min~$=1$ again but
only sampling above the cutoff value of 16 are saved in
the object `MLEbin.MEPS.cutoff16`, obtained by running the next line (which is not evaluated by
this vignette but is run once in `data-raw/simulate-data2.R`):
```{r, eval=FALSE}
MLEbin.MEPS.cutoff16 <- MLEbin.simulate(cut.off = 16)
```

Figure S.37 is then
```{r}
MLEmid.MLEbin.hist(MLEbin.MEPS.cutoff16)
```

Figure S.38 shows the confidence intervals:
```{r}
MLEmid.MLEbin.conf(MLEbin.MEPS.cutoff16)
```

And Table S.5 is:
```{r, echo=TRUE, results='asis'}
knitr::kable(MLEmid.MLEbin.table(MLEbin.MEPS.cutoff16))
```
